Digital Analysis of Paintings

### 2014-04-30 paper.tex
Do paper 11.

\newacronym{APICSS}{APICSS}{Art Painting Image Color Semantics}
\newacronym{RGB}{RGB}{Red, Green, Blue}
\newacronym{HSL}{HSL}{Hue, Saturation, Luminance}
\newacronym{HSV}{HSV}{Hue, Saturation, Value}

@@

Segmentation of brushstrokes is, understandably, difficult - especially
digitally where the image is usually only two dimensional.

>>>>

\newacronym{APICSS}{APICSS}{Art Painting Image Color Semantics}
\newacronym{RGB}{RGB}{Red, Green, Blue}
\newacronym{HSL}{HSL}{Hue, Saturation, Luminance}
\newacronym{HSV}{HSV}{Hue, Saturation, Value}
\newacronym{PCA}{PCA}{Principal Component Analysis}
\newacronym{t-SNE}{t-SNE}{t-Distributed Stochastic Neighbor Embedding}

@@

Segmentation of brushstrokes is, understandably, difficult - especially
digitally where the image is usually only two dimensional, but one can consider
the texture of the painting with more ease and can provide a good insight into
the artists style.

Filters are becoming less popular for painting analysis as the filter often
normalises the image to a certain extent in the process of analysis. Other
approaches like wavelets and pixel-based representations, such as texton
approaches, are becoming more popular to avoid this issue.

It should be noted that whilst textons are a building block of the text of an
image, they are not identical to brushstrokes.

To construct the codebook of textons, 5000 patches are selected in random
locations from each painting in the training set. These patches are then
clustered and the most exemplary patch (the central patch) is selected as the
most representative for that cluster.

From this codebook it is then possible to generate a histogram for any image
which estimates the distribution of these textons across the image. The
histogram is created by applying a sliding window across the painting and
selecting the nearest texton in euclidean space to the window. Finally the
histogram is normalised to sum up to 1.

The authors note that these histograms could be built up for different sizes of
textons to increase the feature space available for analysis, for their own
experiments they use six different scales of texton.

Because the aim of this research was to work closely with experts in the field,
the authors also came up with a method for visualising this information. For an
image processing system this may not be required, but could be used to extract
more useful information from the analysis.

Because of the high dimension space it is necessary to reduce the
dimensionality of the results to allow it to be human readable. Usually this
would be performed by \gls{PCA}, but the way in which the data is structured,
\gls{PCA} is too lossy for texton histograms; it has a linear nature whist the
texton histograms may be high-dimensionally non-linear and it focuses on
preserving the global structure rather than local structure of high-dimensional
points.

There are other techniques which might have been applied to solve these
problems, but the authors found some shortcomings with these techniques which
would have made real-life visualisation difficult.

To cope with this, a new method for dimensionality reduction was invented:
\gls{t-SNE}. This method hinges around keeping
the conditional probabilities between the high and low dimension spaces
similar.

Interestingly, in the experiments \gls{PCA} was first use to reduce the
dimensions down to 50, and then \gls{t-SNE} was
used to reduce down to two dimensions. Suggesting that, for very high dimensions,
\gls{t-SNE} is not very effective.

The results on 117 high-resolution grayscale Van Gogh paintings show that most
of the non-Van Gogh paintings appear on the peripheries of the visualised
two-dimensional graph, apart from two specific forgeries: the Wacker forgery
which had fooled experts for many years, but can successfully be found by
looking at global features. The other, created by Gaugin, may remain undetected
for the same reason.

Visualisation of dated Van Gogh work does show some difference between the two
time periods considered, but not enough to cluster or classify upon with any
degree of accuracy.

% How many clusters?

This approach is definitely a useful one, especially given that the textons are
learned from existing work, making analyses such as trying to date an artists
work from sets of his known work very applicable. Although colour and global
features are not considered by the textons described in this work, colour would
be a simplistic feature to add and global features could be part of a separate
technique which is later included with the histogram.

The paper does leave some answered question, one notable one is how best to
decide the number of clusters to build the texton histogram from and which
scales the textons work best at. For use by experts with \gls{t-SNE} applied
this might be some useful work, but \gls{t-SNE} seem to have limited
application outside this form of data set. A comparison between \gls{PCA} and
\gls{t-SNE} would, perhaps, show the advantage of the latter and an explanation
as to why \gls{PCA} was applied before \gls{t-SNE} on their own experimental
results would help others to better apply such methods correctly.
